% Submission cover letter, updated iter 34 of the overnight loop (2026-05-24).
% Iter-19 drafted the original cover letter; iter 34 adds a paragraph on the
% iter-23--31 Delaunay-extension cluster (honest, partial, conditional) and
% updates the page-count and reader-navigation paragraphs to reflect the
% iter-33 paper-shortening pass and the post-iter-33 raw vs.~effective page
% counts. Stand-alone document, not part of the manuscript proper. Accompanies
% the submission package together with paper_main.pdf and paper_supplement.pdf.
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\setlength{\parskip}{0.5em}
\setlength{\parindent}{0pt}

% Allow URLs to break in long bibliographic-style links.
\hypersetup{breaklinks=true}

\begin{document}

\noindent
Yitzchak Shmalo \\
Einstein Institute of Mathematics \\
Hebrew University of Jerusalem \\
Edmond J.~Safra Campus, Givat Ram, Jerusalem 9190401, Israel \\
\texttt{yitzchak.shmalo@gmail.com}

\vspace{0.6em}
\hfill 2026-05-24

\vspace{1.2em}
\noindent
To the Editors, \\
\emph{[Journal name]}

\vspace{0.8em}
\noindent
Dear Editors,

Please find attached our manuscript ``Boundary and Interface Classes for
Universal Chern Hamiltonians via Doubling and Jamming'' submitted for
consideration. The submission package consists of the main paper
(\texttt{paper\_main.pdf}, 33 pages), an online supplement
(\texttt{paper\_supplement.pdf}, 251 pages raw, of which approximately 160
pages are load-bearing mathematics; the remaining material is internal-%
development audit, framing, navigation, and inventory content that a
math-content reader can skip on first reading and that is explicitly
flagged as such), and a companion code repository linked from the
abstract. A short summary of what the paper does, and a short explanation
of how we are framing it, follows.

\paragraph{Contribution.}
The paper develops a finite-dimensional doubling-and-jamming construction
for boundary and interface classes of Higson--Prodan universal Chern
Hamiltonians on triangulated surfaces with boundary. Given a finite
triangulated oriented surface with boundary, we double it along the boundary,
suppress the reflected copy by a large on-site potential, and study the
resulting jammed Hamiltonian by Schur-complement analysis. The principal
ingredients are: (i) the intrinsic Poincar\'e--Lefschetz boundary
Hamiltonian and its comparison with the original-side compression of the
doubled closed Hamiltonian; (ii) seven Schur-complement estimates for
finite-$M$ jamming (compressed resolvent, first-order effective resolvent,
copy leakage, eigenvalue localization, contour margin, window separation,
and the original--copy coupling support); (iii) a periodic triangular Bloch
certificate, with closed-bulk gap, Chern signs, and integer Toeplitz
interface winding all proved internally on the straight periodic interface;
(iv) internal closed-bulk gap and Chern theorems for the three flat-torus
benchmark scopes (triangular, hexagonal, square), unified by a general
operator-algebraic Chern convergence lemma; (v) an internal Combes--Thomas
exponential decay lemma; and (vi) a complete companion numerical pipeline
that records finite-matrix behavior of all the mechanisms above. These
ingredients are unconditional inside the manuscript: their proofs do not
invoke any external closed-bulk theorem.

\paragraph{Delaunay-extension partial progress.}
Beyond the periodic flat-torus benchmark scopes of ingredient~(iv), the
supplement contains a dedicated Delaunay-extension chapter that develops
quantitative tools for extending the internal closed-bulk analysis from
periodic flat-torus triangulations to the broader quasi-uniform Delaunay
class (every triangle of bounded minimum interior angle and bounded
edge-length ratio, plus the Euclidean Delaunay circumdisk property). The
chapter contains eighteen rigorously proved results, organized into three
explicit tiers: \emph{fourteen unconditional results} (bounded-geometry
parameter lemmas, translation and vertex-perturbation coherence lemmas,
inclusion-coherence and cap-product flip lemmas, single-flip and multi-flip
Schur-complement bounds, common-refinement existence in the union sense,
block-isometry and block-cap-product matrix-element identities, and a
quasi-isometric coherence theorem assembled by cascade); \emph{three
conditional theorems} (a Delaunay-gap stability statement, a
different-dimension coherence theorem, and a three-step composition
theorem) each conditional on a single shared sub-conjecture
(sub-complex refinement existence in the strict inclusion sense); and
\emph{one proved negative result} (an obstruction theorem ruling out the
natural ``refinement-in-the-union-sense'' approach to that sub-conjecture
via an explicit renormalization-mismatch computation, illustrated in the
cleanest possible flat-torus setting). This is partial progress, not a
closure: the Delaunay-extension cluster does \emph{not} eliminate the
bounded-geometry case as an external dependency, and the conditional
theorems remain conditional. What the chapter adds is a quantitatively
explicit framework, with sharp constants and a documented obstruction,
within which a future closure of the sub-conjecture could upgrade three
conditional theorems to unconditional ones. We are presenting this
chapter as honest, mathematically-careful research-medium content
beyond the periodic flat-torus / hexagonal / square benchmark cases of
ingredient~(iv), not as a replacement for the external
bounded-geometry input.

\paragraph{What still depends on Higson--Prodan.}
We are honest about the scope of the unconditional part. For the general
bounded-geometry case, the headline Roe-algebra theorem identifying the
doubled-and-jammed interface class
$\partial_Y\!\bigl(E_+(\pi(P_\pm))\bigr) \in K_1\!\bigl(C^*(Y\subset X)\bigr)$
still uses the Higson--Prodan closed-bulk gap and Chern theorem as an
external input. The bounded-geometry closed-bulk gap and the
bounded-geometry closed-bulk Chern integer are imported as a single named
external hypothesis (the published Higson--Prodan PRL and arXiv articles,
cited in the manuscript), and the manuscript marks every place where this
hypothesis is used. The internal Combes--Thomas decay we prove gives the
locality toolkit that a bounded-geometry gap proof would need, but it does
not by itself supply the gap; the Delaunay-extension chapter described
above gives quantitatively explicit relative-stability tools in the
quasi-uniform Delaunay class but, by the documented obstruction theorem
and the unresolved sub-conjecture, does not by itself close the
bounded-geometry case either. Closing this last step is identified in the
paper as the single research-hard external dependency and as a natural
follow-up project rather than a result of the present work.

\paragraph{Editorial framing of the submission.}
The submitted manuscript is organized along this split. Part~A collects the
unconditional results: the finite-dimensional construction, the seven
Schur-complement estimates, the periodic triangular Bloch certificate and
its Toeplitz winding, the three internal flat-torus benchmark gap and Chern
theorems with the unifying lemma, the internal Combes--Thomas decay, and
the companion numerics. Part~B states the global Roe-algebra interface
theorem in its full bounded-geometry generality and the conditional integer
winding for general partitions; both are conditional on the published
Higson--Prodan PRL result, and we present them as a conditional framework
rather than as unconditional theorems. The supplement contains an explicit
Part~A--Part~B inventory (the iter-17 carve-out table) and a per-figure /
per-table audit confirming that no numerical figure is used as a proof
input and that the Part placement is consistent at the caption level. The
Delaunay-extension chapter sits inside Part~A as additional internal
infrastructure of the unconditional kind (fourteen unconditional results)
together with three conditional theorems that are flagged as conditional
inside the chapter. We chose this two-part presentation deliberately: it
makes the conditional dependency on the published Higson--Prodan PRL result
fully visible and confines it to Part~B, so the unconditional Part~A
material---including the Delaunay-extension unconditional content---can be
evaluated on its own terms.

\paragraph{Why this might be of interest.}
The doubling-and-jamming construction itself is the main technical
ingredient: it converts a closed-bulk universal Chern theorem into
finite-matrix boundary and interface statements via the simplicial double,
with quantitative Schur-complement control of the jamming error. The
three flat-torus benchmark cases give worked closed-form examples in which
the closed-bulk inputs that Higson--Prodan import from the universal model
are proved internally, removing the conditional dependency entirely on
those benchmark families. The periodic triangular case in particular
admits an integer Toeplitz winding on the straight periodic interface, also
proved internally. The Delaunay-extension chapter develops quantitative
quasi-isometric relative-stability tools for the broader bounded-geometry
context, with eighteen rigorously proved results across the three tiers
described above and a single sharply identified remaining research
obstruction. This combination---a quantitative finite-dimensional
boundary framework, plus benchmark cases in which the conditional input is
eliminated, plus a Delaunay-extension cluster of relative-stability tools
with an honest obstruction analysis, plus an explicit cataloguing of what
remains conditional---should be of interest to readers in mathematical
physics and operator algebras working on bulk--interface correspondences
for topological insulators.

\paragraph{Supplement and code.}
The online supplement (251 pages raw) contains the full appendix proofs for
the Schur-complement estimates and the boundary comparison, the eight
internal HP-replacement results (six benchmark closed-bulk gap and Chern
theorems plus the operator-algebraic unification and the Combes--Thomas
decay), the consolidated internal chapter, the HP-dependency inventory,
the Delaunay-extension chapter with its eighteen results across the three
tiers, the topological diagrams, and the numerical record (twenty-five
figures and tables). The supplement also contains internal-development
audit, framing, navigation, and inventory subsections that document the
manuscript's editorial history; these are marked with
``skip-on-first-reading'' notices at their headings, and condensed into a
single ten-row master table at the start of the supplement's reader-%
navigation section. A math-content reader who skips these flagged
subsections encounters an effective supplement of approximately 160 pages
of load-bearing mathematics; a verification-focused or editorial reader
retains full access to the underlying audit trail by consulting the
flagged subsections individually. The companion code repository (URL in
the abstract) contains the data-generation scripts for every numerical
figure and table.

\paragraph{Statements.}
The manuscript has not been published elsewhere and is not under
consideration at another journal. We have no conflicts of interest to
declare.

We hope the editors will find the work suitable for the journal, and we
look forward to the referees' assessment.

\vspace{1.5em}
\noindent Sincerely, \\[1.0em]
Yitzchak Shmalo \\
Einstein Institute of Mathematics \\
Hebrew University of Jerusalem

\end{document}
